---
title: "Example #1"
---

## Introduction

Some examples of using "flexplot" and "lmer" based on explanations from Dustin Fife ([Video Link](https://www.youtube.com/watch?v=eVuQlGDZxHs&list=PL8F480DgtpW9_IT7xN1XeRF_dglZmK0nM&index=2)). Modified with tidyverse packages and plots (ggplot) to enable additional data understanding. My annotations and observations in line with code. Wrapped up in a Quarto package.

## Example #1

```{r}
require(flexplot)
require(lme4)
require(dplyr)
```

### Data about alcohol use in teenagers is distributed with flexplot.

```{r}
data(alcuse)
# see some data
head(alcuse, 10)
# rows
paste0(nrow(alcuse), " rows of data.")
# number of groups and data points
alcuse_sum <- alcuse %>% group_by(ID) %>% 
  summarise(mean = mean(ALCUSE), n_ALCUSE = n()) %>% print(n = 20)
```

```{r}
# number of groups check 
# 3 measures for each ID
nrow(filter(alcuse_sum, n_ALCUSE == 3))

```

##### Comments

There are 3 measurements of ALCUSE for each of 82 IDs.

Probably a boxplot is more useful to get a feel for the data (TODO)

```{r}
summary(alcuse)

```

### Model of alcohol use. No predictors.

```{r}
# Fixed effect + random effects
# Fixed effect: ~1 is gammma 00
# Random effects: 1 is Uij for ID 
mod = lmer(ALCUSE~1 + (1|ID), data=alcuse)
visualize(mod, plot = "model", sample = 20)
summary(mod)

```

##### Comments

Dustin's model visualize() plot does not look like this one despite the code being identical.

The visualize function defaults to 3 random IDs (?). I change it (with sample = n)

Looks like Dustin's data is normalized to 1

Dustin gets Fixed Effect and Random Effects plotted. I just get: object (lmerMod).

However the summary(mod) is identical to Dustin's.

```{r}
# always useful to look this way
summary(mod)
```

##### Comments

The model Intercept 0.9220 is simply the mean of the ALCUSE variable

```{r}
signif(mean(alcuse$ALCUSE),4)

```

### Random intercepts model

```{r}
# random intercepts using AGE_14 as predictor for fixed IDs
# AGE_14 predictor (fixed gradients)
# (1|ID) tells intercepts may vary (random intercepts)
rand.i <- lmer(ALCUSE~1 + AGE_14 + (1|ID), data = alcuse)
# by default 3 IDs are plotted
sample_n <- 6
# if all groups are plotted, one gets funny comments :)
# very nice humour
# sample_n <- 82
visualize(rand.i, plot = "model", sample=sample_n)
summary(rand.i)

```

##### Comments

Fixed effect (Intercept) means 0.65130 drinks on average at age 14.

For subsequent years the number of drinks increases on average by 0.27065.

### Random slopes (gradient) model

```{r}
# unrealistic 
# -1 means intercept is fixed
# need to consider the level 1 and 2 models to specify the syntax correctly
rand.s <- lmer(ALCUSE~1 + AGE_14 + (-1 + AGE_14|ID), data = alcuse)
sample_n <- 10
visualize(rand.s, plot = "model", sample=sample_n)
summary(rand.s)
```

##### Comments

Despite the warning, visualize seems to do the job with 10 samples

### Random slopes and intercepts model

```{r}
# probably what we want
# need to consider the level 1 and 2 models to specify the syntax correctly
rand.is <- lmer(ALCUSE~1 + AGE_14 + (AGE_14|ID), data = alcuse)
sample_n <- 10
visualize(rand.is, plot = "model", sample=sample_n)
summary(rand.is)
```

##### Comments

-   ok
